---
title: "Analysis Process"
author: "draft"
date: "10/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE, dpi=300)

```

```{r}
# Analyze election/parolee data 
# Find local elections where parolee voting rights could swing the vote

library(data.table)
library(dplyr)
library(ggplot2)
library(highcharter)
library(DT)
library(here)

# load election data
elec <- read.csv(here("data/local elections.csv"), stringsAsFactors = FALSE)

# load county parolee estimates
par <- read.csv(here("data/Ratio_Parolees_to_EligibleVoters.csv"), stringsAsFactors = FALSE)
```
  
```{r}

# calc diff btwn winning candidate and losing/2nd place candidate
elec$diff <- elec$winning_vote_count - elec$losing_vote_count

# join parolee tot to county election data
dat <- 
elec %>% 
  filter(office !="Proposition" & !is.na(diff)) %>% 
  left_join(par, by=c("county"="County")) %>% 
  mutate(flag = ifelse(par_tot_est >= diff, "parolee count exceeds margin", "no"),
         flag_wide = ifelse(par_tot_est >= diff - 100, "parolee count within 100", "no"))

# export data
write.csv(dat, here("data/LocalElections_Prop17_Voters.csv"), row.names = FALSE)

```
  
Does the number of parolees living in the county equal or match the number of votes that won the election?
```{r}
table(dat$flag, dat$county)
```


Does the number of parolees living in the county come within 100 votes of the winning number? (i.e. close call, could swing)
```{r}
table(dat$flag_wide, dat$county)
```

### Estimated Number of Parolees Living in Each County  
  
#### County Map
    
```{r}

```
  
    
#### Table   
  
```{r}

datatable(par %>% 
            select(-c(cvap_moe, ratio_par_cvap)) %>% 
            rename("Estimated Number of Parolees" = par_tot_est,
                   "Estimated Number of Eligible Voters" = cvap_est,
                   "Parolees per 100k Eligible Voters" = rate_per_100kcvap),
          rownames = FALSE) %>% 
  formatRound(c(2, 3, 4), 0) 


```
   

### Which elections could be most influeced by parolee voting rights?  
  
* In the data we collected, local school board elections were most frequently the elections that were won by margins smaller than the number of would-be parolee voters.  
  
```{r}
library(kableExtra)

dat %>% 
  filter(flag_wide !="no") %>% 
  group_by(office) %>% 
  summarize(count = length(county)) %>% 
  arrange(-count) %>% 
   kbl() %>%
  kable_minimal(full_width = F)


```
  
### Which elections were won by margins that are less than/close to the number of parolees who could be eligible to vote if Prop 17 passes?  
  
```{r fig.height=9, fig.width=9.5}

dat %>% 
  filter(flag_wide != "no") %>% 
  mutate(name = paste0(office, " ", office_2)) %>% 
  ggplot(aes(x = reorder(name, diff), y = diff, fill=county)) +
  geom_bar(stat="identity") +
  xlab("")+
  ylab("Vote Difference between Winning and Losing Position")+
  coord_flip() +
  theme_minimal()

```
  